---
title: "Homework 1, Question 1"
author: "Nicholas Keeley"
date: "7/21/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1.a

*What is the response variable in this analysis? What is predictor in this analysis?*

The response variable in this analysis would be the "waiting" variable, representing the waiting time until next eruption. This is predicted based off of the predictor variables, "eruptions," which represents Old Faithful's most recent eruption duration in minutes.

## Question 1.b

*Produce a scatterplot of the two variables. How would you describe the relationship between the two variables?*

&nbsp; 
&nbsp; 

```{r read data, echo=FALSE}
# Reading in data.
data <- faithful
attach(data)
```

```{r plot, echo=FALSE}
plot(eruptions, waiting, xlab = "Duration of Last Eruption (minutes)", 
ylab = "Wait Time Until Next Eruption (minutes)", 
main = "Old Faithful Next Eruption Wait Time on Last Eruption Duration\n (minutes)")
```

There appears to be evidence of a linear relationship between eruption time and waiting time. Althought I am concerned about the relative paucity of observations beteween 2.5 minutes and 3.5 minutes (indicating the possibility of clustering), there appears to be a stronger visual cue for assessing these two variables as having a linear relationship.

## Question 1.c

*What is the correlation between the eruption times and waiting times for the next eruption? Interpret this correlation contextually. How reliable is this interpretation?*

```{r 1.c, echo=FALSE}
cor.1 <- cor.test(eruptions, waiting)
cor.1
```

The correlation between eruption time and waiting time is 0.901. This correlation, in conjunction with visual cues from the scatter plot produced earlier, suggests a very strong positive relationship between the variables in question.

## Question 1.d

*Use the lm() function to fit a linear regression for the two variables. Where are the values of Bˆ1, Bˆ0, R2, and sigmaˆ2 for this linear regression?*

```{r 1.d, echo=FALSE}
result <- lm(waiting~eruptions)
summary(result)
```

NOTE: In this document, the " ' " symbol is somtimes used to symbolize a "hat," such as in the estimate of the intercept B0'.\newline

Bˆ1 : 10.7296\newline 
Bˆ0 : 33.4744\newline
R^2 : 0.8115\newline
Variance of the predicted response variable (sigma'^2) : 184.8233

## Question 1.e

*Interpret the values of Bˆ1, Bˆ0 contextually. Does the value of Bˆ0 make sense in this context?*

Bˆ1 is the estimated coefficient of the predictor variable for this linear regression model, and Bˆ0 is the estimated intercept of the linear regression model. In this context, Bˆ1 represents the change in the estimated waiting time for the next eruption for each additional minute of the last eruption duration. Bˆ0 can be interpreted to mean that as eruption duration approaches 0, waiting time approaches 33.47 minutes. However, this value has limited use in this context, since the range of observed X values (eruption duration) does not extend to 0.

## Question 1.f

*Use the anova() function to produce the ANOVA table for this linear regression. What is the value of the ANOVA F statistic? What null and alternative hypothe- ses are being tested here? What is a relevant conclusion based on this ANOVA F statistic?*

```{r 1.f, echo=FALSE}
anova(result)
```

The value of the F-statistic in this ANOVA table is 1162.1. The null hypothesis being tested by this F-statistic is that the coefficient B1, the change in waiting time/change in eruption duration, is equal to 0. This would suggest that a change in eruption duration would have no effect on the waiting time. The alternative hypothesis is that B1 is not equal to zero, and that eruption duration has some effect on waiting time - in either the positive or negative direction. So, this can be considered a two-sided test. Because the F-statistic is formulated by the Sum of Squares of the Regression over the Residual Sum of Squares, this massive number suggests that variance within the linear regression model is primarily attributed to the regression model, implying an effect on the response variable that discounts the null hypothesis. The miniscule p-value within this ANOVA table supports the rejection of the null hypothesis as well. Eruption duration does appear to have some sort of an effect on waiting time.

## Question 1.g

*Obtain the 95% confidence interval for the slope, B1. Is this confidence interval consistent with your conclusion from part 1f? Briefly explain.*

``` {r 1.g, echo=FALSE}
confint(result, level=.95) # 10.10996 11.34932
```

95% confidence interval for slope: 10.10966 <= B1 <= 11.34932

As seen below, manually typing the corresponding values for producing a confidence interval utilizing the mean squares residual produced in the ANOVA table produces approximately the 
same values as the confint() function. The confidence interval is consistent with my conclusion in part 1f.

```{r 1.g.2, echo=TRUE}
# Manual version: estimate +/- multiplier * se
se = 0.3148 # se = sqrt(MSres/n)
est = 10.7296 # from summary(result)
multiplier <- qt(0.975, 270) # 1.968789
est + (multiplier*se) # 11.34937
est - (multiplier*se) # 10.10983
```

## Question 1.h

*The latest eruption at Old Faithful lasted for 3.5 minutes. Obtain an appropriate 95% interval that predicts the waiting time for the next eruption.*
&nbsp; 
&nbsp; 

``` {r 1.h, echo=FALSE}
## 1.h
newdata <- data.frame(eruptions=3.5) # Given X
pre_result<-predict.lm(result, newdata, level=0.95, interval="prediction")
```

```{r 1.h.1, echo=TRUE}
summary(pre_result)
# Lower bound: 59.36332
# Predicted value: 71.02814
# Upper bound: 82.69296
```

95% prediction interval of y at x=3.5: 59.36332 <= y <= 82.69296

## Question 1.i

*What is the 95% interval for the average waiting time for the next eruption among current eruptions that last 3.5 minutes?*

```{r 1.i, echo=FALSE}
## 1.i
conf_result<-predict.lm(result, newdata, level=0.95, interval="confidence")
```

```{r 1.i.1, echo=TRUE}
summary(conf_result)
# Lower bound: 70.32211
# Estimated mean response: 71.02814
# Upper bound: 71.73417
```

95% confidence interval for mean response at x=3.5: 70.32211 =< E(y|x0) =< 71.73417

## Question 1.j
*Create a residual plot, an ACF plot of the residuals, and the QQ plot of the residuals. Based on these plots, are the regression assumptions met? Is your answer surprising, given the context of this data set?*

```{r i.j, echo=FALSE}
## 1.j
plot(result$fitted.values, result$residuals, xlab = "Fitted Values", ylab = "Residuals", main="Residual Plot")
abline(h=0, col="red")
acf(result$residuals, lag.max = 10, main="ACF Plot")
qqnorm(result$residuals)
qqline(result$residuals, col="red")
```

All of the assumptions for a simple linear regression are met except the assumption that error terms aren't correlated. The data points display a general linear pattern, the residuals across the fitted line appear to average to zero, variance within each fitted-value subsection appears evenly spread across the vertical axis, and the residual values appear to be normally distributed. However, according to the ACF plot, error terms with a time point lag of 1 surpassed the designated criical value, indicating a correlation that is statistically significant. Although I was a little suspicious of the slight clustering on either extreme of the scatter plot's X-axis, I had no reliable reason to suspect that the error terms were correlated in anyway. This makes sense since the ACF plot is one of the only tools learned within this module that can calculate correlation between error points. Most of the the other plots reveal information about multiple assumptions, but the ACF plot is singularly unique in its ability to reveal error term correlation.
